Improving real-time trend estimates using local parametrisation of polynomial regression filters

### 2024-03-13 article_jos.qmd
typo + simplify preservation of trends

Fourier decomposition allows us to analyze any time series as a sum of harmonic series, and each component (trend, cycle, seasonal, irregular) is associated with a set of frequencies.
Noting $\omega = 2\pi/p$, the harmonic series of frequency $\omega$ represents a series that repeats itself every $p$ periods.
For example, for a monthly series (12 observations per year), seasonal movements are those that repeat every year: they are therefore associated with frequencies $2\pi/12$ (annual periodicity), $2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5$.
In this paper, we consider that the trend-cycle is associated with frequencies in the interval $[0, 2\pi/12]$, i.e. movements recurring at least every 12 months.
Even if different frequencies are sometimes used (e.g. $[0, 2\pi/36]$ to consider only cycles of at least 36 months), this has no impact on the results of the study.
The other frequencies, $[2\pi/12, \pi]$ are associated with the irregular component (undesirable oscillations).

@@

We have the following properties for the moving average $M_{\boldsymbol\theta}$:

- To preserve a constant trend $X_t=a$ we need:
$$
\forall t:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_ka=a\sum_{k=-p}^{+f}\theta_k=a.
$$
The sum of the coefficients of the moving average $\sum_{k=-p}^{+f}\theta_k$ must then be equal to $1$.

- To preserve linear trends $X_t=at+b$ we need:
$$
\forall t:\:M_{\boldsymbol\theta}(X_t)=\sum_{k=-p}^{+f}\theta_kX_{t+k}=\sum_{k=-p}^{+f}\theta_k[a(t+k)+b]=a\sum_{k=-p}^{+f}k\theta_k+(at+b)\sum_{k=-p}^{+f}\theta_k=at+b.
$$
Which is equivalent to:
$$
\sum_{k=-p}^{+f}\theta_k=1
\quad\text{and}\quad
\sum_{k=-p}^{+f}k\theta_k=0.
$$
- In general, $M_{\boldsymbol\theta}$ preserves polynomials of degree $d$ if and only if:
$$
\sum_{k=-p}^{+f}\theta_k=1
 \text{ and }
\forall j \in \left\llbracket 1,d\right\rrbracket:\:
\sum_{k=-p}^{+f}k^j\theta_k=0.
$$
- If $M_{\boldsymbol\theta}$ is symmetric ($p=f$ and $\theta_{-k} = \theta_k$) and preserves polynomials of degree $2d$ then it also preserves polynomials of degree $2d+1$.

>>>>

Fourier decomposition allows us to analyze any time series as a sum of harmonic series, and each component (trend, cycle, seasonal, irregular) is associated with a set of frequencies.
Noting $\omega = 2\pi/s$, the harmonic series of frequency $\omega$ represents a series that repeats itself every $s$ periods.
For example, for a monthly series ($s=12$ observations per year), seasonal movements are those that repeat every year: they are therefore associated with frequencies $2\pi/12$ (annual periodicity), $2\pi/12\times 2=2\pi/6,\dots,2\pi/12\times 5$.
In this paper, we consider that the trend-cycle is associated with frequencies in the interval $[0, 2\pi/12]$, i.e. movements recurring at least every 12 months.
Even if different frequencies are sometimes used (e.g. $[0, 2\pi/36]$ to consider only cycles of at least 36 months), this has no impact on the results of the study.
The other frequencies, $[2\pi/12, \pi]$ are associated with the irregular component (undesirable oscillations).

@@

For the moving average $M_{\boldsymbol\theta}$, it can be shown that:

- $M_{\boldsymbol\theta}$ preserves polynomials of degree $d$ if and only if:
$$
\sum_{k=-p}^{+f}\theta_k=1
 \text{ and }
\forall j \in \left\llbracket 1,d\right\rrbracket:\:
\sum_{k=-p}^{+f}k^j\theta_k=0.
$$
- If $M_{\boldsymbol\theta}$ is symmetric ($p=f$ and $\theta_{-k} = \theta_k$) and preserves polynomials of degree $2d$ then it also preserves polynomials of degree $2d+1$.
